---
title: "P1 Code"
author: "Audrey Himes"
date: "2/9/2023"
output: html_document
---
```{r}
# loading packages
library(tidyverse)
library(tidytext)
library(ggwordcloud)
library(gutenbergr)
library(textdata)
library(DT)
library(patchwork)
```

## EDA
```{r}
# reading in the data
df0 <- read.csv('/Users/alexaowen/Documents/Fourth Year/Spring 2023/DS 4002/Project 1/REVIEWS.csv')

# remove NA values
df1 <- na.omit(df0)

# keep only relevant variables
df <- df1[,2:3]
str(df)

df <- df %>% mutate(GROUP = case_when(RATING < 4 ~ 'LOW',
  RATING < 8 ~ 'MED',
  RATING < 11 ~ 'HIGH'))

df$GROUP <- as.factor(df$GROUP)
str(df)
```

### Categorizing the Data 
```{r}
# categorize each review as low, medium, high based on the rating
df_LOW <- df %>% filter(RATING == 2 | RATING == 1 | RATING == 3)
df_MED <- df %>% filter(RATING == 4 | RATING == 5 | RATING ==6 | RATING ==7)
df_HIGH <- df %>% filter(RATING == 8 | RATING == 9 | RATING ==10)

table(df$GROUP)
```

### Plot of Ratings by Group
```{r}
plot_explore <- ggplot(data=df, aes(x=RATING, color=GROUP, fill=GROUP)) + 
  geom_bar() +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5)
plot_explore
ggsave("plot_explore.png", plot=plot_explore)
```

## EDA for LOW group
```{r, warning=FALSE}
df_LOW1 <- data.frame(df_LOW$REVIEW)

# breaking the review up into individual 'words'
tidy_LOW_words <- df_LOW1 %>%    
  unnest_tokens(word, df_LOW.REVIEW) %>%
  filter(!grepl('[0-9]', word))

# cleaning out all of the 'stop words'
full_LOW_words <- tidy_LOW_words %>%
  anti_join(stop_words) %>%
  # filter(!grepl('movie', word)) %>%
  # filter(!grepl('film', word)) %>%
  # filter(!grepl('story', word)) %>%
  count(word, sort=TRUE) %>%
  mutate (GROUP = 'LOW')

# data table of word frequencies
datatable(full_LOW_words)

# making a word cloud
cloud_LOW <- ggplot(full_LOW_words[1:50,], aes(label = word, size = n)) +
  geom_text_wordcloud() +
  scale_size_area(max_size=10) +
  theme_minimal()
cloud_LOW
ggsave("cloud_LOW.png", plot=cloud_LOW)
```

## EDA for MED group
```{r}
df_MED1 <- data.frame(df_MED$REVIEW)

# breaking the review up into individual 'words'
tidy_MED_words <- df_MED1 %>%    
  unnest_tokens(word, df_MED.REVIEW) %>%
  filter(!grepl('[0-9]', word))

# cleaning out all of the 'stop words'
full_MED_words <- tidy_MED_words %>%
  anti_join(stop_words) %>%
  # filter(!grepl('movie', word)) %>%
  # filter(!grepl('film', word)) %>%
  # filter(!grepl('story', word)) %>%
  count(word, sort=TRUE)%>%
  mutate (GROUP = 'MED')

# data table of word frequencies
datatable(full_MED_words)

# making a word cloud
cloud_MED <- ggplot(full_MED_words[1:50,], aes(label = word, size = n)) +
  geom_text_wordcloud() +
  scale_size_area(max_size=10) +
  theme_minimal()
cloud_MED
ggsave("cloud_MED.png", plot=cloud_MED)
```

## EDA for HIGH group
```{r}
df_HIGH1 <- data.frame(df_HIGH$REVIEW)

# breaking the review up into individual 'words'
tidy_HIGH_words <- df_HIGH1 %>%    
  unnest_tokens(word, df_HIGH.REVIEW) %>%
  filter(!grepl('[0-9]', word))

# cleaning out all of the 'stop words'
full_HIGH_words <- tidy_HIGH_words %>%
  anti_join(stop_words) %>%
  # filter(!grepl('movie', word)) %>%
  # filter(!grepl('film', word)) %>%
  # filter(!grepl('story', word)) %>%
  count(word, sort=TRUE)%>%
  mutate (GROUP = 'HIGH')

# data table of word frequencies
datatable(full_HIGH_words)

# making a word cloud
cloud_HIGH <- ggplot(full_HIGH_words[1:50,], aes(label = word, size = n)) +
  geom_text_wordcloud() +
  scale_size_area(max_size=10) +
  theme_minimal()
cloud_HIGH
ggsave("cloud_HIGH.png", plot=cloud_HIGH)
```

##AFINN Analysis: Low Group
```{r}
LOW_affin <- full_LOW_words %>%
  inner_join(get_sentiments("afinn"))
LOW_affin

#did weighted mean because each word was used a different number of times
wml <- weighted.mean(LOW_affin$value, LOW_affin$n)
sdl <- sd (LOW_affin$value)
sl <- sum (LOW_affin$value)

#0 is empty because it would just include words with no sentiment making it unnecessary 
ggplot (LOW_affin, aes (x = `value`)) +
  geom_bar(aes (fill = "blue")) + 
  theme_minimal() + 
  scale_x_discrete(limit = c(-5:5)) +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  theme (legend.position = "none") +
  labs (x = 'AFINN Rating', y = 'Count', title = 'Count for Each Ranking from Low Group AFINN Analysis')

```

## AFINN Analysis: Medium Group
```{r}
MED_affin <- full_MED_words %>%
  inner_join(get_sentiments("afinn"))
MED_affin

wmm <- weighted.mean(MED_affin$value, MED_affin$n)
sdm <- sd (MED_affin$value)
sm <- sum (MED_affin$value)

ggplot (MED_affin, aes (x = `value`)) +
  geom_bar(aes (fill = "blue")) + 
  theme_minimal() + 
  scale_x_discrete(limit = c(-5:5)) +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  theme (legend.position = "none") +
  labs (x = 'AFINN Rating', y = 'Count', title = 'Count for Each Ranking from Medium Group AFINN Analysis')
```

## AFINN Analysis: High Group
```{r}
HIGH_affin <- full_HIGH_words %>%
  inner_join(get_sentiments("afinn"))
HIGH_affin

wmh <- weighted.mean(HIGH_affin$value, HIGH_affin$n)
sdh <- sd(HIGH_affin$value)
sh <- sum (HIGH_affin$value)

ggplot (HIGH_affin, aes (x = `value`)) +
  geom_bar (aes (fill = "blue")) + 
  theme_minimal() + 
  scale_x_discrete(limit = c(-5:5)) +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  theme (legend.position = "none") +
  labs (x = 'AFINN Rating', y = 'Count', title = 'Count for Each Ranking from High Group AFINN Analysis')
```
## AFINN Results Table 
```{r}
results_1 <- cbind (wml, sdl, sl)
results_2 <- cbind (wmm, sdm, sm)
results_3 <- cbind (wmh, sdh, sh)
results <- rbind(results_1, results_2, results_3)
colnames(results)<-c("Weighted Mean","Standard Deviation","Sum")
rownames(results) <- c("Low Group", "Medium Group", "High Group" )
results
```

## AFINN Analysis Counts Plot
```{r}
afinn_BIG <- rbind(LOW_affin, MED_affin, HIGH_affin)
afinn_BIG

afinn_BIG$GROUP <- factor(afinn_BIG$GROUP, levels = c("HIGH", "MED", "LOW"))

ggplot(data=afinn_BIG, aes(x= value, fill = GROUP)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(if_else (condition = count>23,
                              as.character(count),""))), position = position_stack(vjust = .5)) +
  labs (x = 'AFINN Analysis Rankings', y = 'Count', title = 'Comparing AFINN Analysis Ranking Counts')
```

## NRC Analysis: Low Group
```{r}
# creating the nrc for low group 
LOW_nrc <- full_LOW_words %>%
  inner_join(get_sentiments("nrc"))
LOW_nrc
table(LOW_nrc$sentiment)

ggplot(data=LOW_nrc, aes(x=sentiment)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  labs (x = 'NRC Sentiment', y = 'Count', title = 'Count for Each Sentiment Grouping from Low Group NRC Analysis')
```

## NRC Analysis: Medium Group
```{r}
# creating the nrc for med group 
MED_nrc <- full_MED_words %>%
  inner_join(get_sentiments("nrc"))
MED_nrc
table(MED_nrc$sentiment)

ggplot(data=MED_nrc, aes(x=sentiment)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  labs (x = 'NRC Sentiment', y = 'Count', title = 'Count for Each Sentiment Grouping from Medium Group NRC Analysis')
```

## NRC Analysis: High Group
```{r}
# creating the nrc for high group 
HIGH_nrc <- full_HIGH_words %>%
  inner_join(get_sentiments("nrc"))
HIGH_nrc
table(HIGH_nrc$sentiment)

ggplot(data=HIGH_nrc, aes(x=sentiment)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  labs (x = 'NRC Sentiment', y = 'Count', title = 'Count for Each Sentiment Grouping from High Group NRC Analysis')
```
## NRC Analysis Counts Plot
```{r}
nrc_BIG <- rbind(LOW_nrc, MED_nrc, HIGH_nrc)
nrc_BIG

nrc_BIG$GROUP <- factor(nrc_BIG$GROUP, levels = c("HIGH", "MED", "LOW"))

ggplot(data=nrc_BIG, aes(x=sentiment, fill = GROUP)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), position = position_stack(vjust = 0.5)) +
  labs (x = 'NRC Sentiment', y = 'Count', title = 'Comparing NRC Analysis Counts')
```

## Bing Analysis: Low Group
```{r}
LOW_bing <- full_LOW_words %>%
  inner_join(get_sentiments("bing"))
LOW_bing

table(LOW_bing$sentiment)

ggplot(data=LOW_bing, aes(x=sentiment)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  labs (x = 'Bing Sentiment', y = 'Count', title = 'Count for Each Sentiment from Low Group Bing Analysis')
```

## Bing Analysis: Medium Group
```{r}
MED_bing <- full_MED_words %>%
  inner_join(get_sentiments("bing"))
MED_bing

table(MED_bing$sentiment)

ggplot(data=MED_bing, aes(x=sentiment)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  labs (x = 'Bing Sentiment', y = 'Count', title = 'Count for Each Sentiment from Medium Group Bing Analysis')
```

## Bing Analysis: High Group
```{r}
HIGH_bing <- full_HIGH_words %>%
  inner_join(get_sentiments("bing"))
HIGH_bing

table(HIGH_bing$sentiment)

ggplot(data=HIGH_bing, aes(x=sentiment)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  labs (x = 'Bing Sentiment', y = 'Count', title = 'Count for Each Sentiment from High Group Bing Analysis')
```

## Bing Analysis Counts Plot
```{r}
bing_BIG <- rbind(LOW_bing, MED_bing, HIGH_bing)
bing_BIG

bing_BIG$GROUP <- factor(bing_BIG$GROUP, levels = c("HIGH", "MED", "LOW"))

ggplot(data=bing_BIG, aes(x=sentiment, fill = GROUP)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), position = position_stack(vjust = 0.5)) +
  labs (x = 'Bing Sentiment', y = 'Count', title = 'Comparing Bing Analysis Counts')
```