---
title: "P1 Code"
author: "Audrey Himes, Alexa Owen, Sharaf Tariq"
date: "2/24/2023"
output: html_document
---
```{r}
# loading packages
library(tidyverse)
library(tidytext)
library(ggwordcloud)
library(textdata)
library(DT)
library(syuzhet)
```

## Initial EDA
```{r}
# reading in the data
df <- read.csv('/Users/audreyhimes/Documents/UVA SP23/DS-4002/DS4002_GroupWork/Project_1/Data/REVIEWS.csv')

# remove NA values
df <- na.omit(df)

# keep only relevant variables
df <- df[,2:3]
str(df)

df <- df %>% mutate(GROUP = case_when(RATING < 4 ~ 'LOW',
  RATING < 8 ~ 'MED',
  RATING < 11 ~ 'HIGH'))

df$GROUP <- as.factor(df$GROUP)
str(df)
```

### Categorizing the Data 
```{r}
# categorize each review as low, medium, high based on the rating
df_LOW <- df %>% filter(RATING == 2 | RATING == 1 | RATING == 3)
df_MED <- df %>% filter(RATING == 4 | RATING == 5 | RATING ==6 | RATING ==7)
df_HIGH <- df %>% filter(RATING == 8 | RATING == 9 | RATING ==10)

table(df$GROUP)
```

### Plot of Ratings by Group
```{r}
df$GROUP <- factor(df$GROUP, levels = c("HIGH", "MED", "LOW"))

plot_explore <- ggplot(data=df, aes(x=RATING, color=GROUP, fill=GROUP)) + 
  geom_bar() +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5)
plot_explore
# ggsave("plot_explore.png", plot=plot_explore)
```

## EDA by Rating Group

### EDA for LOW group
```{r, warning=FALSE}
df_LOW1 <- data.frame(df_LOW$REVIEW)

# breaking the review up into individual 'words'
tidy_LOW_words <- df_LOW1 %>%    
  unnest_tokens(word, df_LOW.REVIEW) %>%
  filter(!grepl('[0-9]', word))

# cleaning out all of the 'stop words'
full_LOW_words <- tidy_LOW_words %>%
  anti_join(stop_words) %>%
  # filter(!grepl('movie', word)) %>%
  # filter(!grepl('film', word)) %>%
  # filter(!grepl('story', word)) %>%
  count(word, sort=TRUE) %>%
  mutate (GROUP = 'LOW')

# data table of word frequencies
datatable(full_LOW_words)

# making a word cloud
cloud_LOW <- ggplot(full_LOW_words[1:50,], aes(label = word, size = n)) +
  geom_text_wordcloud() +
  scale_size_area(max_size=10) +
  theme_minimal()
cloud_LOW
# ggsave("cloud_LOW.png", plot=cloud_LOW)
```

```{r}
# word cloud without "movie", "film", and "story"
clean_LOW_words <- full_LOW_words %>%
  filter(!grepl('movie', word)) %>%
  filter(!grepl('film', word)) %>%
  filter(!grepl('story', word)) %>%
  # count(word, sort=TRUE) %>%
  mutate (GROUP = 'LOW')

# data table of word frequencies
datatable(clean_LOW_words)

# making a clean word cloud
cloud_LOW_clean <- ggplot(clean_LOW_words[1:50,], aes(label = word, size = n)) +
  geom_text_wordcloud() +
  scale_size_area(max_size=10) +
  theme_minimal()
cloud_LOW_clean
```

### EDA for MED group
```{r}
df_MED1 <- data.frame(df_MED$REVIEW)

# breaking the review up into individual 'words'
tidy_MED_words <- df_MED1 %>%    
  unnest_tokens(word, df_MED.REVIEW) %>%
  filter(!grepl('[0-9]', word))

# cleaning out all of the 'stop words'
full_MED_words <- tidy_MED_words %>%
  anti_join(stop_words) %>%
  # filter(!grepl('movie', word)) %>%
  # filter(!grepl('film', word)) %>%
  # filter(!grepl('story', word)) %>%
  count(word, sort=TRUE)%>%
  mutate (GROUP = 'MED')

# data table of word frequencies
datatable(full_MED_words)

# making a word cloud
cloud_MED <- ggplot(full_MED_words[1:50,], aes(label = word, size = n)) +
  geom_text_wordcloud() +
  scale_size_area(max_size=10) +
  theme_minimal()
cloud_MED
# ggsave("cloud_MED.png", plot=cloud_MED)
```

```{r}
# word cloud without "movie", "film", and "story"
clean_MED_words <- full_MED_words %>%
  filter(!grepl('movie', word)) %>%
  filter(!grepl('film', word)) %>%
  filter(!grepl('story', word)) %>%
  # count(word, sort=TRUE) %>%
  mutate (GROUP = 'MED')

# data table of word frequencies
datatable(clean_MED_words)

# making a clean word cloud
cloud_MED_clean <- ggplot(clean_MED_words[1:50,], aes(label = word, size = n)) +
  geom_text_wordcloud() +
  scale_size_area(max_size=9) +
  theme_minimal()
cloud_MED_clean
```


### EDA for HIGH group
```{r}
df_HIGH1 <- data.frame(df_HIGH$REVIEW)

# breaking the review up into individual 'words'
tidy_HIGH_words <- df_HIGH1 %>%    
  unnest_tokens(word, df_HIGH.REVIEW) %>%
  filter(!grepl('[0-9]', word))

# cleaning out all of the 'stop words'
full_HIGH_words <- tidy_HIGH_words %>%
  anti_join(stop_words) %>%
  # filter(!grepl('movie', word)) %>%
  # filter(!grepl('film', word)) %>%
  # filter(!grepl('story', word)) %>%
  count(word, sort=TRUE)%>%
  mutate (GROUP = 'HIGH')

# data table of word frequencies
datatable(full_HIGH_words)

# making a word cloud
cloud_HIGH <- ggplot(full_HIGH_words[1:50,], aes(label = word, size = n)) +
  geom_text_wordcloud() +
  scale_size_area(max_size=10) +
  theme_minimal()
cloud_HIGH
# ggsave("cloud_HIGH.png", plot=cloud_HIGH)
```

```{r}
# word cloud without "movie", "film", and "story"
clean_HIGH_words <- full_HIGH_words %>%
  filter(!grepl('movie', word)) %>%
  filter(!grepl('film', word)) %>%
  filter(!grepl('story', word)) %>%
  # count(word, sort=TRUE) %>%
  mutate (GROUP = 'HIGH')

# data table of word frequencies
datatable(clean_HIGH_words)

# making a clean word cloud
cloud_HIGH_clean <- ggplot(clean_HIGH_words[1:50,], aes(label = word, size = n)) +
  geom_text_wordcloud() +
  scale_size_area(max_size=9) +
  theme_minimal()
cloud_HIGH_clean
```

## AFINN Analysis

### AFINN Analysis: Low Group
```{r}
LOW_afinn <- full_LOW_words %>%
  inner_join(get_sentiments("afinn"))
LOW_afinn

#did weighted mean because each word was used a different number of times
wml <- weighted.mean(LOW_afinn$value, LOW_afinn$n)
sdl <- sd (LOW_afinn$value)
sl <- sum (LOW_afinn$value)

#0 is empty because it would just include words with no sentiment making it unnecessary 
ggplot (LOW_afinn, aes (x = `value`)) +
  geom_bar(aes (fill = "blue")) + 
  theme_minimal() + 
  scale_x_discrete(limit = c(-5:5)) +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  theme (legend.position = "none") +
  labs (x = 'AFINN Rating', y = 'Count', title = 'Count for Each Ranking from Low Group AFINN Analysis')

```

### AFINN Analysis: Medium Group
```{r}
MED_afinn <- full_MED_words %>%
  inner_join(get_sentiments("afinn"))
MED_afinn

wmm <- weighted.mean(MED_afinn$value, MED_afinn$n)
sdm <- sd (MED_afinn$value)
sm <- sum (MED_afinn$value)

ggplot (MED_afinn, aes (x = `value`)) +
  geom_bar(aes (fill = "blue")) + 
  theme_minimal() + 
  scale_x_discrete(limit = c(-5:5)) +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  theme (legend.position = "none") +
  labs (x = 'AFINN Rating', y = 'Count', title = 'Count for Each Ranking from Medium Group AFINN Analysis')
```

### AFINN Analysis: High Group
```{r}
HIGH_afinn <- full_HIGH_words %>%
  inner_join(get_sentiments("afinn"))
HIGH_afinn

wmh <- weighted.mean(HIGH_afinn$value, HIGH_afinn$n)
sdh <- sd(HIGH_afinn$value)
sh <- sum (HIGH_afinn$value)

ggplot (HIGH_afinn, aes (x = `value`)) +
  geom_bar (aes (fill = "blue")) + 
  theme_minimal() + 
  scale_x_discrete(limit = c(-5:5)) +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  theme (legend.position = "none") +
  labs (x = 'AFINN Rating', y = 'Count', title = 'Count for Each Ranking from High Group AFINN Analysis')
```

### AFINN Results Table 
```{r}
afinn_results_1 <- cbind (wml, sdl, sl)
afinn_results_2 <- cbind (wmm, sdm, sm)
afinn_results_3 <- cbind (wmh, sdh, sh)
afinn_results <- rbind(afinn_results_1, afinn_results_2, afinn_results_3)
colnames(afinn_results)<-c("Weighted Mean","Standard Deviation","Sum")
rownames(afinn_results) <- c("Low Group", "Medium Group", "High Group" )
afinn_results
```

### AFINN Analysis Counts Plot
```{r}
afinn_BIG <- rbind(LOW_afinn, MED_afinn, HIGH_afinn)
afinn_BIG

afinn_BIG$GROUP <- factor(afinn_BIG$GROUP, levels = c("HIGH", "MED", "LOW"))

ggplot(data=afinn_BIG, aes(x= value, fill = GROUP)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(if_else (condition = count>23,
                              as.character(count),""))), position = position_stack(vjust = .5)) +
  labs (x = 'AFINN Analysis Rankings', y = 'Count', title = 'Comparing AFINN Analysis Ranking Counts')
```

## NRC Analysis

### NRC Analysis: Low Group
```{r}
# creating the nrc for low group 
LOW_nrc <- full_LOW_words %>%
  inner_join(get_sentiments("nrc"))
LOW_nrc
table(LOW_nrc$sentiment)

LOW_nrc$sentiment <- factor(LOW_nrc$sentiment, levels = c("anger", "disgust", "fear", "sadness", "negative", "positive", "anticipation", "joy", "surprise", "trust"))

ggplot(data=LOW_nrc, aes(x=sentiment)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  labs (x = 'NRC Sentiment', y = 'Count', title = 'Count for Each Sentiment Grouping from Low Group NRC Analysis')
```

### NRC Analysis: Medium Group
```{r}
# creating the nrc for med group 
MED_nrc <- full_MED_words %>%
  inner_join(get_sentiments("nrc"))
MED_nrc
table(MED_nrc$sentiment)

MED_nrc$sentiment <- factor(MED_nrc$sentiment, levels = c("anger", "disgust", "fear", "sadness", "negative", "positive", "anticipation", "joy", "surprise", "trust"))

ggplot(data=MED_nrc, aes(x=sentiment)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  labs (x = 'NRC Sentiment', y = 'Count', title = 'Count for Each Sentiment Grouping from Medium Group NRC Analysis')
```

### NRC Analysis: High Group
```{r}
# creating the nrc for high group 
HIGH_nrc <- full_HIGH_words %>%
  inner_join(get_sentiments("nrc"))
HIGH_nrc
table(HIGH_nrc$sentiment)

HIGH_nrc$sentiment <- factor(HIGH_nrc$sentiment, levels = c("anger", "disgust", "fear", "sadness", "negative", "positive", "anticipation", "joy", "surprise", "trust"))

ggplot(data=HIGH_nrc, aes(x=sentiment)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  labs (x = 'NRC Sentiment', y = 'Count', title = 'Count for Each Sentiment Grouping from High Group NRC Analysis')
```

### NRC Analysis Counts Plot
```{r}
nrc_BIG <- rbind(LOW_nrc, MED_nrc, HIGH_nrc)
nrc_BIG

nrc_BIG$GROUP <- factor(nrc_BIG$GROUP, levels = c("HIGH", "MED", "LOW"))
nrc_BIG$sentiment <- factor(nrc_BIG$sentiment, levels = c("anger", "disgust", "fear", "sadness", "negative", "positive", "anticipation", "joy", "surprise", "trust"))

ggplot(data=nrc_BIG, aes(x=sentiment, fill = GROUP)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), position = position_stack(vjust = 0.5)) +
  labs (x = 'NRC Sentiment', y = 'Count', title = 'Comparing NRC Analysis Counts') +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))
```

## Bing Analysis

### Bing Analysis: Low Group
```{r}
LOW_bing <- full_LOW_words %>%
  inner_join(get_sentiments("bing"))
LOW_bing

table(LOW_bing$sentiment)

ggplot(data=LOW_bing, aes(x=sentiment)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  labs (x = 'Bing Sentiment', y = 'Count', title = 'Count for Each Sentiment from Low Group Bing Analysis')
```

### Bing Analysis: Medium Group
```{r}
MED_bing <- full_MED_words %>%
  inner_join(get_sentiments("bing"))
MED_bing

table(MED_bing$sentiment)

ggplot(data=MED_bing, aes(x=sentiment)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  labs (x = 'Bing Sentiment', y = 'Count', title = 'Count for Each Sentiment from Medium Group Bing Analysis')
```

### Bing Analysis: High Group
```{r}
HIGH_bing <- full_HIGH_words %>%
  inner_join(get_sentiments("bing"))
HIGH_bing

table(HIGH_bing$sentiment)

ggplot(data=HIGH_bing, aes(x=sentiment)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.5) +
  labs (x = 'Bing Sentiment', y = 'Count', title = 'Count for Each Sentiment from High Group Bing Analysis')
```

### Bing Analysis Counts Plot
```{r}
bing_BIG <- rbind(LOW_bing, MED_bing, HIGH_bing)
bing_BIG

bing_BIG$GROUP <- factor(bing_BIG$GROUP, levels = c("HIGH", "MED", "LOW"))

ggplot(data=bing_BIG, aes(x=sentiment, fill = GROUP)) +
  geom_bar() +
  theme_minimal()  +
  geom_text(stat='count', aes(label=after_stat(count)), position = position_stack(vjust = 0.5)) +
  labs (x = 'Bing Sentiment', y = 'Count', title = 'Comparing Bing Analysis Counts')
```

## Sentiment values and vectors
```{r}
vector_LOW_afinn <- get_sentiment(full_LOW_words$word, method = "afinn")
vector_LOW_nrc <- get_sentiment(full_LOW_words$word, method = "nrc")
vector_LOW_bing <- get_sentiment(full_LOW_words$word, method = "bing")

sentiments_LOW <- data.frame(vector_LOW_afinn, vector_LOW_nrc, vector_LOW_bing)
sentiments_LOW

as.vector(c(mean(sentiments_LOW$vector_LOW_afinn), mean(sentiments_LOW$vector_LOW_nrc), mean(sentiments_LOW$vector_LOW_bing)))
```

```{r}
vector_MED_afinn <- get_sentiment(full_MED_words$word, method = "afinn")
vector_MED_nrc <- get_sentiment(full_MED_words$word, method = "nrc")
vector_MED_bing <- get_sentiment(full_MED_words$word, method = "bing")

sentiments_MED <- data.frame(vector_MED_afinn, vector_MED_nrc, vector_MED_bing)
sentiments_MED
```

```{r}
vector_HIGH_afinn <- get_sentiment(full_HIGH_words$word, method = "afinn")
vector_HIGH_nrc <- get_sentiment(full_HIGH_words$word, method = "nrc")
vector_HIGH_bing <- get_sentiment(full_HIGH_words$word, method = "bing")

sentiments_HIGH <- data.frame(vector_HIGH_afinn, vector_HIGH_nrc, vector_HIGH_bing)
sentiments_HIGH
```

```{r}
LOW_total_vector <- as.vector(c(mean(sentiments_LOW$vector_LOW_afinn), mean(sentiments_LOW$vector_LOW_nrc), mean(sentiments_LOW$vector_LOW_bing))) 

MED_total_vector <- as.vector(c(mean(sentiments_MED$vector_MED_afinn), mean(sentiments_MED$vector_MED_nrc), mean(sentiments_MED$vector_MED_bing)))

HIGH_total_vector <- as.vector(c(mean(sentiments_HIGH$vector_HIGH_afinn), mean(sentiments_HIGH$vector_HIGH_nrc), mean(sentiments_HIGH$vector_HIGH_bing)))

values_TOTAL <- data.frame(LOW_total_vector, MED_total_vector, HIGH_total_vector)
rownames(values_TOTAL) <- c("AFINN Score", "NRC Score", "BING Score")

values_TOTAL$Percent_Change <- abs(((values_TOTAL$HIGH_total_vector - values_TOTAL$LOW_total_vector) / values_TOTAL$LOW_total_vector)) * 100
values_TOTAL
```

